---
title: "TFM_Juan_Hernandez"
author: "Juan_Hernandez_Villena"
date: "`r format(Sys.Date(),'%e de %B, %Y')`"
output:
  pdf_document:
    df_print: kable
    toc: yes
    toc_depth: 5
  html_document:
    toc: yes
    toc_depth: 5
    df_print: kable
params: 
  file: lake.csv
editor_options: 
  chunk_output_type: console
---

The _purl_ function allows grouping all the code to be presented in a PDF o similar

```{r purl = F}
#purl("TFM.Rmd", output = "codigo.R")
```

```{r Libraries, warning=FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(naniar)
library(visdat)
library(ggplot2)
library(VIM)
library(MissMech)
library(mice)
library(tidyverse)
library(EnvStats)
library(coin)
```

```{r options}
# Code parameters
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=FALSE,
               prompt=FALSE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
              # include = FALSE,
               fig.align='center')
opts_knit$set(width=75)
```

# Database

## Descriptive analysis

Upload the data. The following data comes from DOI: 10.1016/j.antiviral.2009.11.008
```{r Upload}
vih <- data.frame(read.csv(params$file, sep = ";", dec = ","))
#Dimensions
dim(vih) 
```
Database with `r dim(vih)[1]` observations/rows, and `r dim(vih)[2]` variables.

I choose only those relevant variables for the study.
```{r}
# Variable selection
VIH <- select(vih, Grupo,sexo,edad,tpo_vih_meses,factor_riesgo_total,starts_with("CD4A"),
              starts_with("CargaV"))
# Variable conversion to factor
cols <- c("Grupo","sexo", "factor_riesgo_total")
VIH[cols] <- lapply(VIH[cols], factor)
# Factor levels
levels(VIH$Grupo) <-  c("EFV", "LVP/r")
levels(VIH$sexo)  <- c("Masculino", "Femenino")
levels(VIH$factor_riesgo_total) <- c("ADVP","Heterosexual","Homosexual","Otros")
# Using log10(x+1) to transform "CargaViral_" variable
VIH[11:15] <- VIH %>% select(starts_with("CargaV")) %>% +1 %>%  log10()
```

```{r VIH estructure}
str(VIH)
# Data and missing estructure plot
vis_dat(VIH) + ylab("Observaciones") + guides(fill = guide_legend(title = "Tipo")) +
  scale_fill_hue(labels = c("Factor", "Entero", "Numérico")) + 
  theme(axis.text = element_text(size = 15),
        axis.title.y = element_text(size=rel(1.5)),
        legend.title=element_text(size=15), 
        legend.text=element_text(size=15)) 
```

```{r VIH summary}

# Statistic summary 
summary(VIH)
# Proportions
prop.table(table(VIH$Grupo,VIH$sexo))

prop.table(table(VIH$factor_riesgo_total))
```

```{r boxplot VIH}
par(mfrow = c(1,2))
VIH %>% select(starts_with("CD4")) %>% boxplot(col = "skyblue")
VIH %>% select(starts_with("CargaViral")) %>% boxplot(col = "orange", las = 0)
```

```{r qqplots}
par(mfrow = c(2,5))
VIH[-c(1:5)] %>% apply(2,qqPlot) 
```
## Database sizes

Now, I proceed to get two database size, N/2 y N/3, where N is the original size (from VIH database). It is considered to maintain the same group proportions.

```{r Databases}
#  N/2 and N/3 databases
# Set a seed
set.seed(545)
# slice_sample function require the row numbers for each group

#N/2
VIH2 <- ungroup(VIH %>% group_by(Grupo) %>% slice_sample(n = nrow(VIH)/4))
#N/3
VIH3 <- ungroup(VIH %>% group_by(Grupo) %>% slice_sample(n = nrow(VIH)/6))
```

Check the databases dimensions

```{r, echo=FALSE}
D <- c("VIH","VIH2","VIH3")
N <-c(dim(VIH)[1],dim(VIH2)[1], dim(VIH3)[1])
Nv <- c(dim(VIH)[2],dim(VIH2)[2], dim(VIH3)[2])
M <- c(n_miss(VIH),n_miss(VIH2),n_miss(VIH3))
Mp <- round(c(prop_miss(VIH),prop_miss(VIH2),prop_miss(VIH3)),3)
VIH_DF <- data.frame("DF" = D , "N_obs" = N, "N_var" = Nv, "N_miss" = M, "P_miss" = Mp*100)
kable(VIH_DF)
```

# Missing analysis 
## Descriptive analysis

Database´s missings percentage by variable
```{r missing DBs}

mv_1 <- data.frame(miss_var_summary(VIH))
mv_2 <-data.frame(miss_var_summary(VIH2))
mv_3 <-data.frame(miss_var_summary(VIH3))

mv_12 <- merge(mv_1,mv_2,by = "variable")
mv_total <- merge(mv_12,mv_3,by = "variable")

# sort them by relevant variable 
mv_var <- mv_total %>% select(variable, starts_with("pct"))
colnames(mv_var) <- c("Variable","VIH","VIH2","VIH3")
mv_var <- arrange(mv_var,desc(VIH))
mv_var[-1] <- round(mv_var[-1],2)
kable(mv_var)
```

```{r missings comparison}
mv_var <- arrange(mv_var,Variable)
# CD4A percentage
mv_var_cd4a <- mv_var[6:10,]
mv_cd4 <- mv_var_cd4a %>% mutate(week = seq(0,48,12))
# Carga_viral percentage
mv_var_cv <- mv_var[1:5,]
mv_cv <- mv_var_cv %>% mutate(week = seq(0,48,12))

par(mfrow = c(1,2))
# Plot: CD4A variable missing % by time 
plot(mv_cd4$week,mv_cd4$VIH, type = "b", pch = 18, col = "blue", xlab = "Semanas",
     ylab = "missings (%)", xlim = c(0,48), ylim = c(0,65), main = "Valor absoluto de CD4",
     xaxt = "n", cex.main = 2, cex.lab = 1.5, cex.axis = 1)
lines(mv_cd4$week,mv_cd4$VIH2, type = "b", pch = 20, col = "red")
lines(mv_cd4$week,mv_cd4$VIH3, type = "b", pch = 20, col = "black")
legend("topleft", legend=c("VIH", "VIH2", "VIH3"),
       col=c("blue","red","black"), lty = 1, cex=1.5)
axis(1,at = seq(0,48,12))

# Plot: Carga_viral variable missing % by time
plot(mv_cv$week,mv_cv$VIH, type = "b", pch = 18, col = "blue", xlab = "Semanas",
     ylab = "", xlim = c(0,48), ylim = c(0,65), main = "Log10 Carga viral de VIH", 
     xaxt = "n", cex.main = 2, cex.lab = 1.5, cex.axis = 1)
lines(mv_cv$week,mv_cv$VIH2, type = "b", pch = 20, col = "red")
lines(mv_cv$week,mv_cv$VIH3, type = "b", pch = 20, col = "black")
legend("topleft", legend=c("VIH", "VIH2", "VIH3"),
       col=c("blue","red","black"), lty = 1, cex=1.5)
axis(1,at = seq(0,48,12))
```


## Missing patterns

```{r pattern}
#Patterns plots with the "VIM" package
aggr(VIH, sortVars = T, prop = T, sortCombs = T,  cex.axis = 0.75, combined = T, axes = F)
aggr(VIH2, sortVars = T, prop = T, sortCombs = T,  cex.axis = 0.75, combined = T, axes = F)
aggr(VIH3, sortVars = T, prop = T, sortCombs = T,  cex.axis = 0.75, combined = T, axes = F)
```

Through flux and fluxplot fuctions, It is possible to evaluate which pattern are in the database. For more information, please visit: https://stefvanbuuren.name/fimd/missing-data-pattern.html

```{r influx outflux}
# Tabla de valores influx y outflux de cada variable
flux(VIH)[,1:3]
flux(VIH2)[,1:3]
flux(VIH3)[,1:3]
```

```{r fluxplot}
par(mfrow = c(1,3))
fluxplot(VIH, main = "VIH", labels = FALSE, cex.main = 2, cex.lab = 2, xlab = "")
fluxplot(VIH2, main = "VIH2", labels = FALSE, cex.main = 2, cex.lab = 2, ylab = "")
fluxplot(VIH3, main = "VIH3", labels = FALSE, cex.main = 2, cex.lab = 2, ylab = "",
         xlab = "")
```

The missing pattern is "monotone"

## Missing mechanisms

```{r TestMCAR}
# No normality MCAR test. Otherwise, use Little test.
VIH %>% select("edad",tpo_vih_meses,starts_with("C"))  %>% TestMCARNormality()
```

# Multiple imputation

Fuction to sort dataset variable by time
```{r funcion orden_IM}
# Función para ordenar variables temporales
orden_IM <- function(x){
  select(x,Grupo,sexo,edad,tpo_vih_meses,factor_riesgo_total,ends_with("_0"),
         ends_with("_12"), ends_with("_24"),ends_with("_36"),ends_with("_48"))
}
```

Sorting the variables 
```{r orden BD}
VIH <- orden_IM(VIH)
VIH2 <- orden_IM(VIH2)
VIH3 <- orden_IM(VIH3)
```

Next, I test three differents imputation options, all focus on PMM method

## By default

```{r mice by default}
VIH_SM <- mice(VIH, m=3, maxit = 5, print = FALSE, seed = 457)
VIH_SM2 <- mice(VIH2, m=3, maxit = 5, print = FALSE, seed = 457)
VIH_SM3 <- mice(VIH3, m=3, maxit = 5,  print = FALSE, seed = 457)

# Logged event
VIH_SM3$loggedEvents
# Logged event treatment
prep_SM <- VIH_SM3$pred  # Get the prediction matrix
# Set the predictors
prep_SM[6:14,"factor_riesgo_total"] <- 0
prep_SM["CargaViral_48",c(1:4,7,9,13)] <- 1
# Use the custom prediction matrix
VIH_SM3 <- mice(VIH3, m=3, maxit = 5, print = FALSE, pred = prep_SM,seed = 457)

# Look out the methods used by default
VIH_SM$method
```

```{r Plot_IM}
densityplot(VIH_SM)
stripplot(VIH_SM)
densityplot(VIH_SM2)
stripplot(VIH_SM2)
densityplot(VIH_SM3)
stripplot(VIH_SM3)
```

Add imputed values to each dataset
```{r}
VIH_SIM <- complete(VIH_SM)
VIH_SIM2 <- complete(VIH_SM2)
VIH_SIM3 <- complete(VIH_SM3)
```

## PMM

```{r mice pmm}
# This time we select the method PMM. All variables will treatment by PMM algorithm
VIH_pmm <- mice(VIH, m=3, maxit = 5, method = "pmm", print = FALSE, seed = 457)
VIH_pmm2 <- mice(VIH2, m=3, maxit = 5, method = "pmm", print = FALSE, seed = 457)
VIH_pmm3 <- mice(VIH3, m=3, maxit = 5, method = "pmm", print = FALSE, seed = 457)
# Logged event
VIH_pmm3$loggedEvents
# Logged event treatment
prep3 <- VIH_pmm3$pred
prep3[6:14,"factor_riesgo_total"] <- 0
prep3["CargaViral_48", c(1:4,7,9,13)] <- 1

VIH_pmm3 <- mice(VIH3, m=3, maxit = 5, method = "pmm", print = FALSE, 
                 pred = prep3,seed = 457)
```

```{r plot_IM_pmm}
densityplot(VIH_pmm)
stripplot(VIH_pmm)
densityplot(VIH_pmm2)
stripplot(VIH_pmm2)
densityplot(VIH_pmm3)
stripplot(VIH_pmm3)
```

```{r}
VIH_IM <- complete(VIH_pmm)
VIH_IM2 <- complete(VIH_pmm2)
VIH_IM3 <- complete(VIH_pmm3)
```

## Midastouch 

```{r mice midas}
method <- c("","logreg","midastouch","midastouch","polyreg","midastouch","midastouch",
            "midastouch","midastouch","midastouch","midastouch","midastouch","midastouch",
            "midastouch","midastouch")

VIH_midas <- mice(VIH, m = 3, method = method, print = FALSE, seed = 457)
VIH_midas2 <- mice(VIH2, m = 3, method = method, print = FALSE, seed = 457)
VIH_midas3 <- mice(VIH3, m = 3, method = method, print = FALSE, seed = 457)
# Logged event
VIH_midas3$loggedEvents
# Logged event treatment 
premi3 <- VIH_midas3$pred
premi3[6:15,"factor_riesgo_total"] <- 0
premi3["CargaViral_48", c(1:4,7,9,13)] <- 1

VIH_midas3 <- mice(VIH3, m = 3,method = method, print = FALSE, pred = premi3, seed = 457)
```


```{r plot_IM_midas}
densityplot(VIH_midas)
stripplot(VIH_midas)
densityplot(VIH_midas2)
stripplot(VIH_midas2)
densityplot(VIH_midas3)
stripplot(VIH_midas3)
```

```{r}
VIH_IMM <- complete(VIH_midas)
VIH_IMM2 <- complete(VIH_midas2)
VIH_IMM3 <- complete(VIH_midas3)
```


# Results

We reproduce the analize made by the dataset's publication, to compare if the results are the same or not.

Use the Mann-Whitney test to compare the variable "group"
 
```{r funcion test}
test <- function(variable,grupo){

  x <- wilcox_test(variable ~ grupo,conf.int = TRUE, distribution = "exact")
  # Valor p
  p <- pvalue(x)
  # Estadístico/valor Z
  z <- statistic(x)
  # Extraemos ambos valores redondeados
  c <- round(c(p,z),3)
  return(c)
}
```

### CD4A
```{r}
# By default
U_SIM_CD <- test(VIH_SIM$CD4A_48,VIH_SIM$Grupo)
U_SIM2_CD <- test(VIH_SIM2$CD4A_48,VIH_SIM2$Grupo)
U_SIM3_CD <- test(VIH_SIM3$CD4A_48,VIH_SIM3$Grupo)
# PMM
U_IM_CD <- test(VIH_IM$CD4A_48,VIH_IM$Grupo)
U_IM2_CD <- test(VIH_IM2$CD4A_48,VIH_IM2$Grupo)
U_IM3_CD <- test(VIH_IM3$CD4A_48,VIH_IM3$Grupo)
# Midastouch
U_IMM_CD <- test(VIH_IMM$CD4A_48,VIH_IMM$Grupo)
U_IMM2_CD <- test(VIH_IMM2$CD4A_48,VIH_IMM2$Grupo)
U_IMM3_CD <- test(VIH_IMM3$CD4A_48,VIH_IMM3$Grupo)
```

### Carga_viral
```{r CargaViral_48 Variable}
# By default
U_SIM_CV <- test(VIH_SIM$CargaViral_48,VIH_SIM$Grupo)
U_SIM2_CV <- test(VIH_SIM2$CargaViral_48,VIH_SIM2$Grupo)
U_SIM3_CV <- test(VIH_SIM3$CargaViral_48,VIH_SIM3$Grupo)
# PMM
U_IM_CV <- test(VIH_IM$CargaViral_48,VIH_IM$Grupo)
U_IM2_CV <- test(VIH_IM2$CargaViral_48,VIH_IM2$Grupo)
U_IM3_CV <- test(VIH_IM3$CargaViral_48,VIH_IM3$Grupo)
# Midastouch
U_IMM_CV <- test(VIH_IMM$CargaViral_48,VIH_IMM$Grupo)
U_IMM2_CV <- test(VIH_IMM2$CargaViral_48,VIH_IMM2$Grupo)
U_IMM3_CV <- test(VIH_IMM3$CargaViral_48,VIH_IMM3$Grupo)
```

```{r}
res_cd <- data.frame(t(data.frame("VIH_defecto" = U_SIM_CD, "VIH2_defecto" = U_SIM2_CD,
                                  "VIH3_defecto" = U_SIM3_CD, "VIH_PMM" = U_IM_CD, 
                                  "VIH2_PMM" = U_IM2_CD, "VIH3_PMM" = U_IM3_CD, 
                                  "VIH_Midas" = U_IMM_CD, "VIH2_Midas" = U_IMM2_CD,
                                  "VIH3_Midas" = U_IMM3_CD)))
                     
res_cv <- data.frame(t(data.frame("VIH_defecto" = U_SIM_CV, "VIH2_defecto" = U_SIM2_CV,
                                  "VIH3_defecto" = U_SIM3_CV, "VIH_PMM" = U_IM_CV, 
                                  "VIH2_PMM" = U_IM2_CV, "VIH3_PMM" = U_IM3_CV, 
                                  "VIH_Midas" = U_IMM_CV, "VIH2_Midas" = U_IMM2_CV,
                                  "VIH3_Midas" = U_IMM3_CV)))

res <- cbind(res_cd,res_cv)
colnames(res) <- c("P_valor(CD4A)","Z(CD4A)","P_valor(log10 CargaViral)", 
                   "Z(Log10 CargaViral)")
kable(res)
```
